---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# blake3

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

`blake3` is an interface to the [BLAKE3](https://github.com/BLAKE3-team/BLAKE3) cryptographic hash function.

To quote blake3's readme:

> BLAKE3 is a cryptographic hash function that is:
> 
> * Much faster than MD5, SHA-1, SHA-2, SHA-3, and BLAKE2.
> * Secure, unlike MD5 and SHA-1. And secure against length extension, unlike SHA-2.
> * Highly parallelizable across any number of threads and SIMD lanes, because it's a Merkle > tree on the inside.
> * Capable of verified streaming and incremental updates, again because it's a Merkle tree.
> * A PRF, MAC, KDF, and XOF, as well as a regular hash.
> * One algorithm with no variants, which is fast on x86-64 and also on smaller architectures.


## Installation

``` r
remotes::install_github("dirkschumacher/blake3")
```

Also make sure to have a very recent rust compiler. The best way to install the rust toolchain is to use [rustup](https://rustup.rs/).

## Example

```{r example}
library(blake3)

input <- "This is a string"
hash <- blake3_hash_raw(charToRaw(input))
(sodium::bin2hex(hash))

```

## License

MIT
